---
title: "COVID-19 Guided Project"
author: "Mona Lee"
date: "2025-05-18"
output: html_document
---

```{r, include=FALSE}
# Best practice: do not install packages in .Rmd
# This document requires the 'tidyverse' package.
# Install it once using: install.packages("tidyverse")
library(tidyverse) #includes dplyr, ggplot2, readr, tibble, etc. knitr::opts_chunk\$set(echo = TRUE)
```

# COVID-19 Guided Project

## Introduction

This project is to explore the efficacy of public health governments according to their message delivery during the COVID-19 pandemic. We will begin by finding the provinces which had the best testing and recovery rates. This will be then followed by exploring the approach the top 5 countries had with their public health response strategies early in the pandemic.

The goal of this project is to determine if there is any pattern in the public health measures that achieved the best testing and recovery rates.

We will begin by briefly evaluating the data by gathering its dimensions, column names, peeking at the table, and viewing a "glimpse" summary.

```{r covid-setup, include=FALSE}
covid_df <- read_csv("C:/Users/monam/OneDrive/Documents/covid19.csv", col_names = TRUE)

dim(covid_df)
colnames(covid_df)
head(covid_df)
glimpse(covid_df)
```

## Data Clean-up

Given that the data set groups certain provinces together for some countries, yet specifies the province for certain countries such as the U.S., we will filter our data to be country-wide (removing state-specific rows and removing the Province-level information).

```{r clean-up 1, echo=FALSE}
covid_df_all_states <- covid_df |>
  filter(Province_State == "All States") |>
  select( - Province_State)
covid_df_all_states
```

To manage the differences involved with measurements that are cumulative vs. daily information, we will further isolate for daily data.

```{r clean-up 2, echo=FALSE}
covid_df_all_states_daily <- covid_df_all_states |>
  select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)
head(covid_df_all_states_daily)
```

## Finding the Top 10 Countries in the Number of Tested Cases

We can begin with getting oversight into how we can get the overall number of people tested, positive results, active, and hospitalized cased by country, reported daily.

```{r overall, echo=FALSE}
covid_df_all_states_daily_sum <-   covid_df_all_states_daily |> 
  group_by(Country_Region) |>
  summarize(tested = sum(daily_tested),
            positive = sum(daily_positive),
            active = sum(positive),
            hospitalized = sum(hospitalizedCurr)
            ) |>
  arrange(desc(tested), by_group = Country_Region)
covid_df_all_states_daily_sum

covid_top_10 <- head(covid_df_all_states_daily_sum, 10)
```

## Finding the Top 5 Countries with the Most Positive Cases Against Tested Cases

Now we can explore: Which 5 countries have had the highest number of positive cases against the number of tests?

```{r vectors & matrices, echo=FALSE}
countries <- covid_top_10$Country_Region
positive_cases <- covid_top_10$positive
tested_cases <- covid_top_10$tested
active_cases <- covid_top_10$active
hospitalized_cases <- covid_top_10$hospitalized

names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
tested_cases
positive_cases
active_cases
hospitalized_cases
```

Now we can identify the top 5 positive against tested cases.

```{r cases, echo=FALSE}
positive_cases/tested_cases
positive_tested_top_5 <- c("United Kingdom"=0.113260617,"United States"=0.108618191,"Turkey"=0.080711720, "Italy"=0.061523368, "Peru"=0.060910738)
positive_tested_top_5
```

We can make a matrix for the relevant values of the top 5 countries for easy display.

```{r covid-matrix, echo=FALSE}
united_kingdom <- c(0.11, 1473672, 166909, 166909, 0)
united_states <- c(0.10, 17282363, 1877179, 1877179, 0)
turkey <- c(0.08, 2031192, 163941, 163941, 0)
italy <- c(0.06, 4091291, 251710, 251710, 1699003)
peru <- c(0.06, 976790, 59497, 59497, 0)

covid_mat1 <- rbind(united_kingdom, united_states, turkey, italy, peru)
colnames(covid_mat1) <- c("Ratio", "Tested", "Positive", "Active", "Hospitalized")
```

```{r chunk-name}
covid_mat1
```
Now, the guided project prompts questions relating the countries' COVID-19 case data to their population size. However the population size isn't provided in the first dataset. Population data can be found at: https://www.kaggle.com/datasets/koryto/countryinfo
```{R population, include=FALSE}
# question3 <- "Which country had the best testing effort relative to its population?"
# 
# question4 <- "Which country was the most affected relative to its population?"
# 
# question5 <- "Which country was the least affected relative to its population?"
```


Now we can report all our findings into a list of the data we can use to answer the following questions: 
* Which top ten countries had the most COVID-19 cases? 
* Which countries had the highest number of positive cases relative to the number of tests? 
* Which country had the best testing effort relative to its population? 
* Which country was the most affected relative to its population? 
* Which country was the least affected relative to its population?


```{r final findings list, echo=FALSE}
question2 <- "Which countries have had the highest number of positive cases against the number of tests?"

answer2 <- c("Positive tested cases" = positive_tested_top_5)

dataframes <- list(covid_df, covid_df_all_states, covid_df_all_states_daily, covid_top_10)
matrices <- list(covid_mat)
vectors <- list(vector_cols, countries)
data_structure_list <- list("dataframes", dataframes, "matrices", matrices, "vectors", vectors)

covid_analysis_list <- list(question2, answer2, data_structure_list)

covid_analysis_list[2]
```

Now as promised, the final part of our exploration is to find out what the top 5 countries had done in their public health response strategies early in the pandemic, and evaluate if there is any pattern in the public health measures that achieved the best testing and recovery rates.
